Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-08-23 paper.qmd
Regenerated Figure 2

projection <- cbind(
    c(-0.2782,0.2156,0.1410,0.8068,0.0804,-0.0315,0.4449),
    c(-0.6090,0.3014,-0.1654,-0.3381,-0.0964,0.5900,0.1979))

@@

scurve_proj1_true_model <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
      geom_point(
        size = 0.8,
        alpha = 0.5,
        color = "#000000") +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#3182bd") + 
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 5) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70")

@@

projection <- cbind(
    c(-0.6296,-0.1452,0.3314,0.2259,-0.4488,-0.0053,-0.4693),
    c(0.5314,0.4634,0.1431,0.2129,-0.0107,0.1477,-0.6443))

scurve_proj2_true_model <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
      geom_point(
        size = 0.8,
        alpha = 0.5,
        color = "#000000") +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#3182bd") + 
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 5) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70")

```{r}
#| label: s-curve-true-model-proj3
#| eval: false

projection <- cbind(
    c(0.11484,0.06997,-0.13638,0.01096,0.01533,0.16257,-0.13615),
    c(0.07805,0.01652,0.17174,-0.07343,0.16177,0.11112,0.04729))

projected <- as.matrix(training_data_scurve |> select(-ID)) %*% projection

projected_true <- as.matrix(true_model_df |> select(-ID)) %*% projection

limits <- 1.5
rng <- range(projected)
projected <- projected/max(abs(rng))
colnames(projected) <- c("P1", "P2")
projected <- data.frame(projected)
obs_labels <- as.character(1:nrow(training_data_scurve))

axis_scale <- 2 * limits/3
axis_pos <- -2/3 * limits

axis_labels <- colnames(training_data_scurve |> select(-ID))
rownames(axes) <- axis_labels

# theta <- seq(0, 2 * pi, length = 50)
# circle <- data.frame(c1 = adj(cos(theta)), c2 = adj(sin(theta)))

scurve_proj3_true_model <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
      geom_point(
        size = 0.8,
        alpha = 0.8,
        color = "#000000") +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#969696") + 
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50")

```{r}
#| label: fig-scurve-true-sc
#| fig-cap: "Two views of the true model (blue lines) in $2\\text{-}D$ projections from $7\\text{-}D$, for the S-curve data (black points). The data is spread along the S shape, and does not vary much from this curve. (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/I5GL23vLiw0>)." 
#| fig-width: 10
#| fig-height: 6
#| fig-pos: H

>>>>

projection <- cbind(
    c(0.45181,0.04747,0.02613,0.25735,0.20622,0.08755,-0.11488),
    c(0.03231,0.23556,0.35166,-0.11741,0.12730,0.09544,0.34262))

axis_scale <- limits/6
axis_pos <- -2/3 * limits

# axis_scale <- limits/4
# axis_pos <- -0.5 * limits

scurve_proj1_true_model <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
      geom_point(
        size = 0.8,
        alpha = 0.5,
        color = "#000000") +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#3182bd") + 
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed()

projection <- cbind(
    c(0.03627,0.15991,0.39768,-0.21032,-0.14886,-0.17589,0.23305),
    c(-0.34731,-0.08183,0.15819,0.31926,0.08105,0.10083,0.25627))

@@

scurve_proj2_true_model <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
      geom_point(
        size = 0.8,
        alpha = 0.5,
        color = "#000000") +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#3182bd") + 
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed()

@@

```{r}
#| label: fig-scurve-true-sc
#| fig-cap: "Two views of the true model (blue lines) in $2\\text{-}D$ projections from $7\\text{-}D$, for the S-curve data (black points). The data is spread along the S shape, and does not vary much from this curve. (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/I5GL23vLiw0>)." 
#| fig-pos: H
#| fig-width: 10
#| fig-height: 5
